# Practical Machine Learning: Course Project

## Loading and preprocessing the data

The data for this project come from this source: [Human Activity Recognition](http://groupware.les.inf.puc-rio.br/har)
Read the train data from the file [Train Data](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)
```{r, echo=TRUE,cache=TRUE}
fileUrl1<-"http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
download.file(fileUrl1,destfile="pml-training.csv")
Training<-read.csv("pml-training.csv")
```
Read the test data from the file [Test data](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)
```{r, echo=TRUE,cache=TRUE}
fileUrl2<-"http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(fileUrl2,destfile="pml-testing.csv")
Testing<-read.csv("pml-testing.csv")
```
Remove the columns where almost all the values is NA, the empty columns and the columns that does not have information about moviment (from 1 to 7) at the training data and the test data
```{r,echo=TRUE,cache=TRUE}
NewTrain<-Training[,colSums(!is.na(Training))>19216]
NewTrain2<-NewTrain[,!apply(NewTrain,2,function(x) any(x==""))]
NewTrain2<-NewTrain2[,-c(1:7)]
Testing<-Testing[,colSums(!is.na(Training))>19216]
Testing<-Testing[,!apply(NewTrain,2,function(x) any(x==""))]
Testing<-Testing[,-c(1:7)]
Testing<-Testing[,-53] #Remove the last column that contens the problem_id
```
Create the cross validation (20% of data) from the training set.
```{r,echo=TRUE}
library(caret)
inTrain<-createDataPartition(y=NewTrain2$classe,p=0.8,list=FALSE)
mytrain<-NewTrain2[inTrain,]
mycv<-NewTrain2[-inTrain,]
```


##Create the Random Forest
Here we create the random forest model
```{r,echo=TRUE}
library(randomForest)
set.seed(124)
RF<-randomForest(classe~ .,data=mytrain,na.action=na.fail)
```
Test the random forest with the cross validation data
```{r,echo=TRUE}
pred1<-predict(RF,mycv);mycv$predRight<-pred1==mycv$classe
table(pred1,mycv$predRight)
pro<-mean(pred1==mycv$classe)*100
er<-100-pro
confusionMatrix(pred1,mycv$classe)
```
The model has a `r pro`% the accuracy with the cross validation data and `r er`% the error. So the cross validation data validates the model.

##Results of the Test set with the Random Forest model

```{r,echo=TRUE}
Pred<-predict(RF,newdata=Testing)
Pred
```
Create the files for the prediction
```{r,echo=TRUE,eval=FALSE}
answers = as.character(Pred)
pml_write_files = function(x){
        n = length(x)
        for(i in 1:n){
                filename = paste0("problem_id_",i,".txt")
                write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
        }
}
pml_write_files(answers)
```



